---
title: "Practical Machine Learning Course Project Report"
author: "Rafael Sulino"
date: "11/08/2020"
output:
  pdf_document: default
  html_document: default
---

## Introduction  

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data Preprocessing  
```{r, cache = TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
### Downloading Data

```{r, cache = TRUE}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="libcurl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="libcurl")
}
``` 

### Reading Data

Read the two csv files into two data frames.

```{r, cache = TRUE}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
Training dataset:

- 19622 observations
- 160 variables

Testing dataset:

- 20 observations
- 160 variables

### Cleaning data


```{r, cache = TRUE}
sum(complete.cases(trainRaw))
```
Removing missing values.

```{r, cache = TRUE}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  

Removing some columns (not used).

```{r, cache = TRUE}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

Training dataset (cleaned):

- 19622 observations
- 53 variables

Testing dataset (cleaned):

- 20 observations
- 54 variables

### Spliting Data

- Training = 70%
- Validation = 30%


```{r, cache = TRUE}
set.seed(1234)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Modeling Data

Random Forest, 5-fold cross validation.


```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
Model performance.

```{r, cache = TRUE}
testdData <- testData[complete.cases(testData),]
```

```{r, cache = TRUE}
predictRf <- predict(modelRf, testData)
confusionMatrix(predictRf, as.factor(testData$classe))
```

```{r, cache = TRUE}
accuracy <- postResample(predictRf, as.factor(testData$classe))
accuracy
oose <- 1 - as.numeric(confusionMatrix(as.factor(testData$classe), predictRf)$overall[1])
oose
```

## Predicting for Test dataset

Apply the model to the original testing dataset.

```{r, cache = TRUE}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Appendix: Figures

1. Correlation Matrix

```{r, cache = TRUE}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```

2. Decision Tree

```{r, cache = TRUE}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```